Supplemental Material for ``Assessing United States county-level exposure for research on tropical cyclones and human health"

### 2020-09-03 ehp_revision/response_to_reviewers.Rmd

---
title: "Response to reviewers' comments for *\"Assessing United States county-level exposure for research on tropical cyclones and human health\"*"
output: bookdown::pdf_document2
toc: false
bibliography: writing/hurr_exposure.bib
header-includes:
- \usepackage{xcolor}
- \usepackage{framed}
- \usepackage{bm}
- \usepackage{subcaption}
- \usepackage{gensymb}
- \renewcommand\familydefault{\sfdefault}
- \usepackage{sansmath}
- \usepackage{bm}
- \usepackage{soul}
- \sansmath
- \usepackage{booktabs}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{pdflscape}
- \usepackage{float}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{fullpage}
- \usepackage{pdflscape}
- \usepackage{tcolorbox}
- \tcbuselibrary{skins,breakable}
- \usepackage{amsmath}
- \usepackage{xtab}
- \usepackage{rotating}
- \allowdisplaybreaks
- \usepackage[markers,nolists]{endfloat}
- \pagenumbering{gobble}
---

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd

---
title: "Response to reviewers' comments for *\"Assessing United States county-level exposure for research on tropical cyclones and human health\"*"
output: bookdown::pdf_document2
toc: false
bibliography: writing/hurr_exposure.bib
header-includes:
- \usepackage{xcolor}
- \usepackage{framed}
- \usepackage{bm}
- \usepackage{subcaption}
- \usepackage{gensymb}
- \renewcommand\familydefault{\sfdefault}
- \usepackage{sansmath}
- \usepackage{bm}
- \usepackage{soul}
- \sansmath
- \usepackage{booktabs}
- \usepackage{array}
- \usepackage{multirow}
- \usepackage{pdflscape}
- \usepackage{float}
- \usepackage{tabu}
- \usepackage{threeparttable}
- \usepackage{threeparttablex}
- \usepackage[normalem]{ulem}
- \usepackage{makecell}
- \usepackage{fullpage}
- \usepackage{pdflscape}
- \usepackage{tcolorbox}
- \tcbuselibrary{skins,breakable}
- \usepackage{amsmath}
- \usepackage{xtab}
- \usepackage{rotating}
- \allowdisplaybreaks
- \usepackage[nomarkers]{endfloat}
- \pagenumbering{gobble}
---

@@

...

In terms of using FEMA declarations as a source for this, however, there are 
some limitations. These declarations are on average tied to the amount of 
damage and other impacts experienced following a storm, but there are a number
of other considerations that drive whether a disaster is declared or not. 
These contributing factors could vary by location (especially from one state
to another) and also over time, especially over the long period we've included
for our other data. Given that, these may not be the best data source to use
to try to capture estimates of the damage or impacts to a community in a multi-year, 
multi-site dataset like the one we present.

FEMA's disaster declarations are issued to provide assistance, based on damage
assessment, to individuals or the entire public in certain locations where
catastrophes overwhelm the state or local government [@mccarthy2014fema], and
any use of them for exposure assessment is secondary. Due to the political
nature of disaster declaration process, these declarations are often subject to
many political and economic factors [@mccarthy2014fema; @logue1981research]. One
epidemiological study specifically focused on exposure assessment for tropical
cyclone epidemiology and compared use of FEMA declarations with other methods of
exposure assessment, including one based directly on the storm hazard of wind
[@grabich2015measuring]. They found that exposure assessment based on FEMA
declarations tended to overassign counties to be "exposed," resulting in false
positives [@grabich2015measuring]. A further concern is that, given the
political and economic nature of these declarations, there is likely variation
across time and geography in the likelihood of a given exposure resulting in a
disaster declaration. This is particularly worrisome for a study that seeks to
explore risk across multiple years and affected communities.

Other sources of data may be better suited to use as a marker of damage to
assess that pathway of indirect health impacts from a tropical cyclone. For
example, data on storm impacts, including property damage or insurance claims
data, are sometimes available through storm databases or publications (e.g.,
NOAA's Storm Events data, *Monthly Weather Review*). While such data could
potentially be used to create an exposure metric, tropical storm damage data
requires some normalization, to incorporate both changes in dollar value over
time and also changes in development in at-risk areas, to be comparable over
extended time periods [@pielke1998].

@@

This question could be explored to some degree through a study that uses the 
physical hazard data provided in this dataset as the exposure and FEMA disaster
declarations as the outcome. Such a study could clarify the degree to which 
each hazard---and its intensity---helps in explaining that probablity that 
a disaster declaration is issued. It could also clarify the degree to which 
other factors, beyond the intensity of these hazards, drive the probability of
a county receiving a disaster declaration. For example, if two counties 
are exposed to a storm-associated hazard of the same intensity, they may 
still have different probabilities of associated societal impacts, as a 
community that is regularly exposed to that hazard may be more "hardened" 
against it. Other differences in infrastructure and factors associated with 
vulnerability could similarly introduce variability.

At a more complex level, studies could explore the causal pathways from the 
initial physical hazard to resulting damage/property loss through to associations
with health outcomes, to provide a clearer idea of the direct and indirect 
pathways leading to health impacts from tropical cyclones.

For both of these studies, however, there may be better sources of damage
estimates to use than FEMA disaster declarations. FEMA disaster declarations 
can be prone to political and other influences, and since their primary function
is not to record damage but rather to be a step in getting help to communities, 
they can be problematic to use as an unbiased estimate of damage levels in 
epidmiological research. One study, for example, found ... . There are other
estimates of property damage that might be less problematic, including the 
estimates recorded in NOAA's Storm Events database and data from private 
insurers, although the quality of disaster loss data is well known to be
less than ideal for secondary research [can be prone to bias through a number
of mechanisms] [ref to Gall paper?].

FEMA declaration is largely subjective, requiring state governors and/or tribal
chiefs to submit a proposal for government aid on the basis that local resources
are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as
exposure classification to hurricanes restricts exposure to 1) those states
whose resources are overwhelmed and 2) those states who apply for FEMA aid.
(FEMA is likely less objective given societal effects (politics) [@kunkel1999])
As a result, other severely impacted counties and states that may not fit FEMA
criteria are excluded and may therefore be incorrectly identified as unexposed,
preventing researchers from assessing the full impact of a given hurricane
[JMF-- the Grabich 2016 study actually found that FEMA assigned the highest
exposure, and may overestimate the health effects of a hurricane; GBA-- let's
think about why that might have had "false positives". I think that would be an
interesting point to include here.]. Researchers have found heterogeneity in
exposure assignments of counties, both within and between storms, when different
metrics of exposure are used [@grabich2015measuring; @grabich2015]. Such
missclassification can have important consequences in assessments of the public
health and economic impacts of tropical storms.

One study used FEMA disaster declarations as a binary indicator of tropical
storm exposure [@grabich2015measuring]. They found that this metric tended to assess a
lot more counties as "exposed" to a storm than distance- or wind-based metrics
[@grabich2015measuring], although they only compared the two exposure metrics as applied
to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
2004).

Data on storm impacts, including property damage or insurance claims data, are
sometimes available through storm databases or publications (e.g., NOAA's Storm
Events data, *Monthly Weather Review*). While such data could potentially be
used to create an exposure metric, tropical storm damage data requires some
normalization, to incorporate both changes in dollar value over time and also
changes in development in at-risk areas, to be comparable over extended time
periods [@pielke1998].

@@

\begin{quotation}
\noindent
\textit{"For another set of tropical cyclones (e.g., Ernesto in 2006, and Bertha
in 1996, Isabel in 2003), there was moderate to good agreement for pairwise
combinations of distance, rain, and wind, but poor agreement for other
combinations of metrics, while for another set of storms (e.g., Matthew in 2004
and Katrina in 2005), there was moderate to good agreement between distance and
rain."}
\end{quotation}

\end{shaded}

**Response:**

The reviewer makes a good point that it is important for us to clarify our
motivation for measuring agreement between the hazard-based metrics, as well as
to show readers how our results from these measures can be used to help in
designing studies and planning statistical analysis for tropical cyclone
epidemiology.

A tropical cyclone brings a number of physical hazards, including the four 
included in our study---wind, rain, flooding, and tornadoes. These separate 
hazards may each cause health impacts and in some cases might act together 
in synergy in elevating health risks. For example, severe wind and rain 
together might create a higher probability of power outages, since the rain 
could soften the ground and make it more likely that the wind can uproot 
trees. These physical hazards tend to follow different patterns, and this is
something that's intuitively understood by atmospheric scientists who study 
the storms. Storm winds tend to decrease rapidly once a storm's center is 
over land, for example, because the storm loses its main energy source---warm
water. Rainfall can persist, and might be particularly severe when the storm 
hits landscape that is not smooth, like the Appalachian Mountains. Tornadoes
can be generated in outer bands of the storm, far from the storm's center, 
but this tends to only happen in the southern half of the US. Rain and flooding
will sometimes align spatially, but you can also have severe flooding downstream
from the area that receives the most severe rain, particularly in coastal areas
with large networks of waterways.

None of these patterns would be particularly surprising to an atmospheric scientist, 
but many epidemiologists might not be aware of them, or of how they play into
exposure assessment for tropical cyclone epidemiology. Many environmental 
epidemiologists have a training that was focused more on epidemiology and 
biostatistics and rely on collaborations with experts in engineering,
atmospheric science, or environmental health to ensure that they understand
and are appropriately applying exposure assessment. In the case of tropical 
cyclones, it is clear that they generate a mixture of physical hazards. 
There has been a lot of methodological development and thought recently in 
environmental epidemiology about how to study mixtures of chemicals---for example, 
the mixture of air pollutants that a person might be exposed to, rather than 
studying the individual effects of single air pollutants. 
These mixture methods require an understanding of how the different components
are associated with each other. In the case of binary, rather than continuous,
exposure, this will be expressed as agreement among the different exposures.

First, if exposure to a pair of different hazards strongly agrees (i.e., when a
county is exposed to one hazard, they are typically also exposed to a second),
this would have several implications for epidemiological research, especially
for multi-year studies:

- From a statistical modeling point of view, if you tried to include each hazard
as a separate independent variable, you would run into problems with
collinearity if fitting the study data with a generalized linear model or other
regression model. One implication of this is that the confidence intervals on
effect estimates for the two variables would be inflated, potentially a lot,
because the model would struggle to figure out which of the two variables the
weight should be placed on, and so this uncertainty would be reflected in the
size of the associated confidence intervals.
- If you see an effect, it would be very hard to try to determine if the causal
pathway goes through the first hazard or the second. To mitigate effects, it is
important to understand the pathway through which they happen, and it might be
easier to get clues towards pathways if the hazards are better separated. By
analogy, if you were able to conduct a controlled experiment to look at the
effects of each hazard, jointly and in combination, you would design separation
in the assignment of these treatments, so you could distinguish between the two
factors. For tropical cyclone epidemiology, we are limited to an observational
analysis, but the area of causal inference is doing a lot to explore how we can
design observational studies and analyze data to come closer to what we might be
able to see with a randomized control study. To determine which of those methods
could be used for tropical cyclone epidemiology, it is important to understand
the interplay of these patterns of exposures.
- If this is the case, and a study is conducted on a single hazard of the storm, 
then there is the risk that results have been confounded by a different hazard
of the storm. 
- On the positive side, in this case the researcher may be able to fit a simpler
statistical model to estimate the joint effects of exposure to the storm as a
whole. Statistical models of mixtures can be complex to lay out, and their 
results can be hard to interpret and explain to others. If all hazards of a 
tropical cyclone follow very similar spatial patterns, then it would be 
reasonable to model tropical cyclone exposure based on one of the hazards
(and it agreement is strong enough, it wouldn't make much difference which 
of the hazards was used for this). This would allow for the use of very 
straightforward statistical modeling, drawing on well-developed models currently
used in environmental epidemiology to study things like the health risk associated
with exposure to a single pollutant or a single physical hazard like outdoor 
temperature.

Second, if a pair of hazards tend to have little spatial overlap among 
counties, this also brings some implications for epidemiological study design
and statistical analysis:

- In this case, a single-hazard analysis could miss a lot of the risk created
by the storm as a whole. This is a concern because there have been some cases
of studies that have assessed exposure based on a single storm, typically wind.
[Grabich example, Darren Sun example]. We are aware of one other paper 
currently under review that takes this same approach of focusing on the wind
hazard of these storms. .... In interpreting these papers, it is helpful to
know if this hazard can stand as a reasonable proxy to the collection of 
physical hazards of the storm. If not, the paper must be interpreted as 
investigating

In the past, much of tropical cyclone epidemiology focused on cases studies of
single storms. Multi-year studies are becoming more common. Both types of
studies are important in exploring both typical and unusual health risks that
can be caused by these storms. Therefore, it's important to understand the
patterns between these different hazards both for single storms and across many
storms. In our analysis, we find that these hazards typically are not in strong
agreement across most storms, suggesting that mixture methods could be
meaningfully applied in most cases, and that there may often be the chance to
separate the role of different hazards through such models. However, we also
find that there are a few storms for which the hazards follow very similar
spatial patterns (Hurricane Floyd in 1999, for example), and in these cases
there may be problems of multi-collinearity and non-separability of different
hazard effects if a researcher tries to fit a model that incorporates each
separate hazard.

Issues:

1. **Interpretation** Effect of one hazard, or storm as a whole?

2. **Confounding**

3. **Multi-collinearity**

4. **Disentangling role of different hazards**

5. **Statistical power** (to separate effects of different hazards)

------------------------------------------------------------------------------

2. These results will not be surprising to an atmospheric scientist / disaster
researcher, but there are environmental epidemiologists who have done research
on other exposures and are now starting to study tropical cyclones (session at
ISEE, for example). It is useful for epidemiologists entering into tropical
cyclone research to understand these facets of tropical cyclone exposure,
including the distinction between the very coastal wind exposures and other
hazards that tend to reach further inland.
3. There are epidemiological studies that have used only one hazard for exposure
assessment. This has most often been the case for wind. If this exposure
assessment is meant to capture "tropical cyclone exposure" in general, with all
the hazards that are related to tropical cyclones, then there is the chance for
exposure misclassification if this single hazard is used.
4. In the past, it was a very complex issue to try to pull together data on all
these hazards. We agree that now that we have provided a dataset with all of
them, epidemiologists should be thinking about multi-hazard analyses. However,
there still will be some barriers to that. [Recent work on multi-pollutant
exposures and the complexities associated with that.] If one of the hazards
dominates in explaining associated health risk, then in some cases it might be
reasonable to focus on that hazard, but in that case it is critical, based on
our results, to remember that the exposure assessment is not capturing all
hazards of the storms.

The largest-scale county-level exposure study is likely that of @zandbergen2009,
which estimated exposure in US counties to all US landfalling Atlantic basin
tropical storms between 1851 and 2003, using both distance and an exposure
metric that incorporated distance and windspeed. They used this exposure
evaluation to create maps of total exposure to tropical storms within US
counties, as well as to explore associations between a county's long-term
exposure to tropical storms and its location, distance from the coast, size, and
shape [@zandbergen2009]. @kruk2010 explored exposure to hurricane-related winds
in the United States, including inland areas, for 1900-2008.

@@

> **Text added to the revised manuscript (relevant additions in bold):**
"The tornado observations from this dataset---**along with the derived version
of the data available through the Storm Prediction Center's National Tornado
Database [@center2020storm]---are the official tornado event data** for the US.
**There are no other collections of tornado data comparable in temporal or
geographical scope,** and so we did not further validate the tornado event
data."

\begin{shaded}
\textbf{R3 Comment 6:}
Table 1: I suggest providing the rationale for picking the thresholds for
classifying continuous exposures into binary ones within the table. 
\end{shaded}

**Response:**

For all outcomes, there are various reasonable choices that can be picked for
the threshold in defining exposure. There is not room to go into these details
within Table 1, so instead we added a separate table to the Supplemental
Material that provides both some reasoning behind the thresholds selected in
this study and also other reasonable thresholds that could be considered.
This is Table S1 of the revised manuscript and has been reproduced as 
Table [x] at the end of this response document.

\input{tables/thresholds_table.tex}

Ultimately, this choice will depend on what a researcher considers as plausible
causal pathways for the outcome he or she is studying. For example, if power
outages form a plausible step in the pathway, then the threshold for gale-force
or strong gale-force winds might be reasonable, while if the pathway is more
likely to be through larger-scale building destruction, a higher threshold would
be more reasonable. Some studies have done sensitivity analyses to consider how
results changed using different exposure metrics. For example, @czajkowski2011
performed a sensitivity analysis using different assumptions about wind decay
away from the center of a storm, while @currie2013, who used a distance-based
exposure metric, considered different buffer distances as a sensitivity
analysis.

@@

- **Wind:** This threshold is the threshold for gale-force and higher winds
based on the Beaufort wind scale. This value is used by the National Hurricane
Center [?] as the outer limit of winds for its wind radii product in storm 
tracking data. Other reasonable choices for a threshold of wind exposure 
could include the two more intense thresholds also used in the NHC [?] wind
radii estimates of storm extent (50 knots and 64 knots, the later of which 
is the threshold for hurricane-force winds on the Beaufort Scale) or other
thresholds on the Beaufort Scale (e.g., "strong gale", which starts at 
41 knots; "storm", which starts at 48 knots; "violent storm", which starts
at 56 knots).

- **Rain:** This threshold is ... Other reasonable thresholds might be ... or
might incorporate different ranges of days around the storm's closest approach
to a county in calculating the cumulative precipitation. For example, ... For
the most part, using our window (two days before to one day after a storm's
closest approach) captured most of the storm-associated precipitation that would
have been counted using a wider window from five days before to two days after a
storm's closest approach, even for a large, slow storm. However, for a storm
that loops and so passes some counties twice, such as Allison (2001), this
four-day window may not be adequate, and lead to an underestimation of total
rainfall associated with the storm relative to estimates for faster-moving
storms. Other storms that looped over or near land and so passed close to US
counties twice include Allison (1989), Alberto (1994), Gordon (1994), Dennis
(1999), Edouard (2002), Fay (2002), Alex (2004), Ivan (2004), Dennis (2005), and
Ophelia (2005). One difficulty in measuring rain exposure consistently across
many different storms is how many days to include in the rain total. Some
storms, especially ones associated with dangerous rainfalls, can move very
slowly, meaning a window of several days needs to be used to fully capture
storm-related rainfall. For example, one study used a period of over four days
to provide storm rainfall totals from the slow-moving Hurricane Danny in 1997
[@medlin2007]. Single-storm studies often measure rainfall as the total rainfall
over the full "event" (e.g., @medlin2007); however, the timing of the event is
determine in a customized way for each storm, making it difficult to extend to
multi-storm analyses. We included more days prior to the storm because
storm-related rains often precede the passing of the storm's center [ref].

@@

been used in previous research in assessing tropical cyclone exposure using
distance from the storm's track as a proxy. Other distance-based exposure
metrics have defined "exposure" to occur only when a storm's center track passes
through a county's boundary (e.g., @zandbergen2009, who describe this metric of
exposure as a "hit" on the county; @kinney2008 is another example that

@@

considered highest exposures for counties directly on the storm's path). This
way of measuring exposure can exclude nearby counties that suffered extreme
conditions from the storm but were not directly on the hurricane's track
[@kruk2010], especially since hurricanes can have a width ([typical storm width;

@@

good place to start. Suggest 30-60km width of the eye of the storm itself] that
is much larger than a typical county's width and since a hurricane's maximum

### 2020-09-09 ehp_revision/response_to_reviewers.Rmd

winds can extend well away from the storm's center. Further, exposure assessed
using this method is associated with the size [@zandbergen2009; @kruk2010] and
shape [@zandbergen2009] of a county. (While some studies have avoided this
problem by determining exposure for equal-area grid cells (e.g., [@kruk2010]),

@@

contributing to the necessity in understanding how such parameters influence a
county's exposure status. In assessing the risk of a given storm based on
distance, recent studies have defined the distance from the storm track affected
by a given storm differently. @czajkowski2011 assessed county-level risk and
exposure based on a three-tiered definition, with primary counties being those

@@

closest to the storm track on either side, secondary counties being adjacent to
primary counties, and tertiary counties adjacent to secondary counties. Such a

@@

definition resulted in an exposure defenition based on an average distance
radius of 120 km [75 miles] on either side of the storm track [@czajkowski2011].

@@

Such a distance is slightly larger than that commonly used by public health
departments (i.e., 100 km) [@czajkowski2011]. Other studies used distance
buffers of 30 kilometers [@currie2013] or of 30, 60, and 100 kilometers as
thresholds for identifying distance-based exposure [@grabich2015measuring] [check the

@@

the county?; JMF-- I think any part of the county]. Another study used a
distance criterion of 60 kilometers from the storm track, classifying all
counties for which the 60 km buffer fully or partially passed through the
county's boundaries as exposed [@grabich2015] [again, check to see if this was

@@

(120 kilometers) from the storm track for a county to qualify as "exposed" to a
storm [@czajkowski2011]. The study found that about 30 direct hurricane
fatalities over their study period (1970--2007) occurred outside of this
distance range from the storm center [@czajkowski2011], suggesting that this

### 2020-09-09 ehp_revision/writing/ehp_discussion.tex
Work some more on response

\textbf{This dataset is limited to the contiguous US. The precipitation data is
extracted and aggregated to the county level from a re-analysis data product.
While the re-analysis product used here (NLDAS-2) only covers the contiguous
US, other similar re-analysis products, as well as other types of precipitation
datasets, have global coverage \parencite{sun2018review}. The wind data are
generated based on a model that is currently US-focused
\parencite{stormwindmodel}, but could be extended to other areas, although this
would require adding new land/sea masks within the associated software, as well
as accounting for differences across storm basins in wind averaging periods
\parencite{harper2010guidelines} and in the direction of cyclonic winds in the
Northern versus Southern Hemisphere. Further, since the core of the wind model
was developed based on data from Atlantic-basin storms
\parencite{willoughby2006parametric}, an extension to other areas should
include separate validation and calibration to ensure it performs appropriately
in those settings. Other wind field modeling software is also available that
already provides a global coverage, like Geoscience Australia's Tropical
Cyclone Risk Model (http://geoscienceaustralia.github.io/tcrm/). For tornado
and flood events, the data described in this paper drew on a US-focused storm
events database, and so international extension of data on these hazards would
require access to similar databases covering other countries. Finally, the
relevant geopolitical boundaries to use for aggregation would vary by country
(e.g., municipalities in Mexico, districts in India).}

@@

Second, these data are aggregated to the county level. \textbf{This spatial
scale allows for easy integration with health outcome data aggregated at the
county level. Such aggregated or ecological data is often used for disaster
epidemiology, as aggregated data may be easier to access than individual-level
data, especially at a scale that covers many locations and a long time period
and so allows higher statistical power and a broader range of exposure levels
within the study \parencite{wakefield2008overcoming}.}

>>>>

\textbf{This dataset is limited to the contiguous US. The precipitation data is
extracted and aggregated to the county level from a re-analysis data product.
While the re-analysis product used here (NLDAS-2) only covers the contiguous
US, other similar re-analysis products, as well as other types of precipitation
datasets, have global coverage \parencite{sun2018review}. The wind data are
generated based on a model that is currently US-focused
\parencite{stormwindmodel}, but could be extended to other areas, although this
would require adding new land/sea masks within the associated software, as well
as accounting for differences across storm basins in wind averaging periods
\parencite{harper2010guidelines} and in the direction of cyclonic winds in the
Northern versus Southern Hemisphere. Further, since the core of the wind model
was developed based on data from Atlantic-basin storms
\parencite{willoughby2006parametric}, an extension to other areas should
include separate validation and calibration to ensure it performs appropriately
in those settings. Alternatively, other wind field modeling software is
available that provides a global coverage, like Geoscience Australia's Tropical
Cyclone Risk Model (http://geoscienceaustralia.github.io/tcrm/). For tornado
and flood events, the data described in this paper drew on a US-focused storm
events database, and so international extension of data on these hazards would
require access to similar databases covering other countries. Finally, the
relevant geopolitical boundaries to use for aggregation would vary by country
(e.g., municipalities in Mexico, districts in India).}

\textbf{The dataset we present focuses on the physical hazards of a tropical
cyclone.  However, health impacts will often some through indirect pathways,
including through damage to property and infrastructure. In future work, it
would be useful to expand this dataset to add data related to these pathways,
to allow research exploring the role of indirect pathways from tropical cyclone
phyiscal hazards to health risk. This expansion could include adding data on
normalized storm-associated damages, or proxy measurements of such damages,
from sources like the US NOAA's Storm Events database and the US Federal
Emergency Management Agency's county-level disaster declarations.}

@@

Second, these data are aggregated to the county level. \textbf{This spatial
scale allows for easy integration with health outcome data aggregated at the
county level. Such data is often used for disaster epidemiology, as aggregated
data may be easier to access than individual-level data, especially at a scale
that covers many locations and a long time period and so allows higher
statistical power and a broader range of exposure levels within the study
\parencite{wakefield2008overcoming}.}

### 2020-09-09 ehp_revision/writing/ehp_response_darlings.tex
Work some more on response

>>>>

# R1 Comment 3

In this paper, we present a dataset that focuses on exposure to the primary
physical hazards of wind, rain, tornadoes, and flooding. We do not include data
related to intermediate pathways, but such data could be a very interesting
future expansion of the data.

There are two elements that are important related to this question. First, many
of the health impacts of tropical cyclones will likely come through indirect
pathways, including through damage to property and infrastructure. Ultimately,
it would be useful to have large-scale databases of these secondary exposure
pathways to use for more complex studies of the pathways between the physical
hazards of these storms and how they affect human health. In this sense,
something like the FEMA declarations could be considered not as a proxy for
assessing exposure to physical hazards, but rather as a proxy of damage caused
by the storm, and so a measurement informative for one of these indirect
pathways from physical hazards of the storm to health impacts, through something
like a mediation analysis. We have added text to the manuscript on this point:

In terms of using FEMA declarations as a source for this, however, there are 
some limitations. These declarations are on average tied to the amount of 
damage and other impacts experienced following a storm, but there are a number
of other considerations that drive whether a disaster is declared or not. 
These contributing factors could vary by location (especially from one state
to another) and also over time, especially over the long period we've included
for our other data. Given that, these may not be the best data source to use
to try to capture estimates of the damage or impacts to a community in a multi-year, 
multi-site dataset like the one we present.

FEMA's disaster declarations are issued to provide assistance, based on damage
assessment, to individuals or the entire public in certain locations where
catastrophes overwhelm the state or local government [@mccarthy2014fema], and
any use of them for exposure assessment is secondary. Due to the political
nature of disaster declaration process, these declarations are often subject to
many political and economic factors [@mccarthy2014fema; @logue1981research]. One
epidemiological study specifically focused on exposure assessment for tropical
cyclone epidemiology and compared use of FEMA declarations with other methods of
exposure assessment, including one based directly on the storm hazard of wind
[@grabich2015measuring]. They found that exposure assessment based on FEMA
declarations tended to overassign counties to be "exposed," resulting in false
positives [@grabich2015measuring]. A further concern is that, given the
political and economic nature of these declarations, there is likely variation
across time and geography in the likelihood of a given exposure resulting in a
disaster declaration. This is particularly worrisome for a study that seeks to
explore risk across multiple years and affected communities.

Other sources of data may be better suited to use as a marker of damage to
assess that pathway of indirect health impacts from a tropical cyclone. For
example, data on storm impacts, including property damage or insurance claims
data, are sometimes available through storm databases or publications (e.g.,
NOAA's Storm Events data, *Monthly Weather Review*). While such data could
potentially be used to create an exposure metric, tropical storm damage data
requires some normalization, to incorporate both changes in dollar value over
time and also changes in development in at-risk areas, to be comparable over
extended time periods [@pielke1998].

However, evidence from prior tropical cyclone epidemiology that doing so is
problematic and should be avoided [@grabich2015measuring].

We have added some discussion of this point...

- Future research could expand the work by [@grabich2015measuring] to determine
the extent of these issues with use of FEMA disaster declarations for
epidemiological exposure assessment. This would need to go beyond the comparison
with exposure assessment based on storm hazards, because sometimes the
characteristics of the affected community modifies the amount of disaster that
results from hazards of a certain severity. For example, a community that it
exposed often might be more "hardened" against damage from the storm (any work
from Seth on this) and so might experience less infrastructure damage from a
tropical cyclone with certain characteristics compared to a community that is
rarely exposed to tropical cyclones. One interesting direction would be to
investigate if there is evidence of differences in probablity of disaster
declarations at geopolitical boundaries, like when comparing counties on either
side of a state border.

There are some cases where damage reports are directly used to infer the
intensity of the weather event. For example, damage surveying is used following
a tornado to assign the tornado's class on the [enhanced] Fujita scale. However,
this is not the intent of FEMA damage reports, which instead are used to
determine post-disaster aid [?] and are influenced by a number of political
factors.

> "Because of the historic lack of direct measurements and remotely sensed
tornado wind speeds at or near group level, damage surveying has remained the
most common form for indicating tornado strength. ... The occurrence of a direct
tornado strict upon a fixed, sufficiently sturdy, and well-calibrated wind
measuring station is quite rare. Only 31 direct in situ tornado observations are
evident between 1894 and 2011." [@edwards2013tornado]

This question could be explored to some degree through a study that uses the 
physical hazard data provided in this dataset as the exposure and FEMA disaster
declarations as the outcome. Such a study could clarify the degree to which 
each hazard---and its intensity---helps in explaining that probablity that 
a disaster declaration is issued. It could also clarify the degree to which 
other factors, beyond the intensity of these hazards, drive the probability of
a county receiving a disaster declaration. For example, if two counties 
are exposed to a storm-associated hazard of the same intensity, they may 
still have different probabilities of associated societal impacts, as a 
community that is regularly exposed to that hazard may be more "hardened" 
against it. Other differences in infrastructure and factors associated with 
vulnerability could similarly introduce variability.

At a more complex level, studies could explore the causal pathways from the 
initial physical hazard to resulting damage/property loss through to associations
with health outcomes, to provide a clearer idea of the direct and indirect 
pathways leading to health impacts from tropical cyclones.

For both of these studies, however, there may be better sources of damage
estimates to use than FEMA disaster declarations. FEMA disaster declarations 
can be prone to political and other influences, and since their primary function
is not to record damage but rather to be a step in getting help to communities, 
they can be problematic to use as an unbiased estimate of damage levels in 
epidmiological research. One study, for example, found ... . There are other
estimates of property damage that might be less problematic, including the 
estimates recorded in NOAA's Storm Events database and data from private 
insurers, although the quality of disaster loss data is well known to be
less than ideal for secondary research [can be prone to bias through a number
of mechanisms] [ref to Gall paper?].

FEMA declaration is largely subjective, requiring state governors and/or tribal
chiefs to submit a proposal for government aid on the basis that local resources
are overwhelmed [@FEMA]. As a result, using FEMA disaster declarations as
exposure classification to hurricanes restricts exposure to 1) those states
whose resources are overwhelmed and 2) those states who apply for FEMA aid.
(FEMA is likely less objective given societal effects (politics) [@kunkel1999])
As a result, other severely impacted counties and states that may not fit FEMA
criteria are excluded and may therefore be incorrectly identified as unexposed,
preventing researchers from assessing the full impact of a given hurricane
[JMF-- the Grabich 2016 study actually found that FEMA assigned the highest
exposure, and may overestimate the health effects of a hurricane; GBA-- let's
think about why that might have had "false positives". I think that would be an
interesting point to include here.]. Researchers have found heterogeneity in
exposure assignments of counties, both within and between storms, when different
metrics of exposure are used [@grabich2015measuring; @grabich2015]. Such
missclassification can have important consequences in assessments of the public
health and economic impacts of tropical storms.

One study used FEMA disaster declarations as a binary indicator of tropical
storm exposure [@grabich2015measuring]. They found that this metric tended to assess a
lot more counties as "exposed" to a storm than distance- or wind-based metrics
[@grabich2015measuring], although they only compared the two exposure metrics as applied
to four storms (Hurricanes Charley, Frances, Ivan, and Jeanne in Florida in
2004).

Data on storm impacts, including property damage or insurance claims data, are
sometimes available through storm databases or publications (e.g., NOAA's Storm
Events data, *Monthly Weather Review*). While such data could potentially be
used to create an exposure metric, tropical storm damage data requires some
normalization, to incorporate both changes in dollar value over time and also
changes in development in at-risk areas, to be comparable over extended time
periods [@pielke1998].
